### 2023-04-09 main/chapter_experiment.tex
add pile-up

% TODO pile-up
\TODO{Distribution in the mean number of primary vertices per crossing, where the
variability in that mean arises from differing beam conditions over run-time.}
Figure~\ref{fig:experiment_run_2_mu}

@@

>>>>

% pile-up
From a theoretical standpoint, we would like to study individual proton
scattering events. But from a practical standpoint, to get enough interesting
data with limited resources we must increase the luminosity.
We therefore expect many proton scatterings per bunch crossing, and hope that
one of them is interesting enough to record.
On reconstruction, each proton-proton scattering in a bunch crossing produces
a vertex.%
\footnote{%
Sometimes each vertex is described as a primary vertex, to distinguish it from
secondary vertices from decaying products. Here, we reserve `primary' for the
main hard scatter of interest in an event.%
}
The number of vertices in a bunch crossing is the pile-up number.
The primary vertex is the main interaction of interest.
It is typically the hardest hard scatter, and is identified in \atlas\ as the
vertex with the largest sum of squared transverse momenta of its
tracks~\cite{atlas2022searches}.

@@

% pile-up
Pile-up interactions cause tracks and other detector hits, which are noise
atop our measurements of the primary vertex.
If not removed, these pile-up signals can skew measurements and create an
illusion missing transverse momentum.
Although precise measurements can help to remove this noise, it reduces the
information content of our data and cannot be cleaned entirely.
In Run~2 data, the number of pile-up vertices per bunch crossing varies around
several tens of vertices, as displayed in
Figure~\ref{fig:experiment_run_2_mu}~\cite{atlas_public_luminosity}.
Note that the variability in the mean pile-up count arises from differing beam
conditions over run-time; for example, the instantaneous luminosity can only
decrease during each LHC beam fill, and the expected pile-up rate decreases
with it.

Having reviewed the LHC on which our detectors are constructed, we now move on
to the design of the \atlas\ detector itself.
